---
title: "Run simulated epidemic history"
author: "Stephane"
date: "3/27/2017"
header-includes:
   - \usepackage{multirow, booktabs, threeparttable}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Run simulated epidemic history

```{r read}
library(knitr)
read_chunk('model0.R')
```

Load packages
```{r libs, warning=FALSE, message=FALSE}
```

Source C functions
```{r source C}

```

Define input parameters
```{r parameters}
```
\input{tab.tex}

Note: incidence and diagnosis rate scaling factors are a priori. Now, there are fitted 
doi:10.1371/journal.pone.0055312.g002 (Fig 2.A)

every individuals start infection at EHI stage


prRecipMat:
  - prob that recipient get infection, conditionning on
	- prob of being risk level 1 (80%) vs risk level 2 (20%)
	- EHI stage (only those recipient get infection)
	- care status (only undiagnosed get infection)
       	- age assortativity (power of age class difference) 
  - intervenes in F matrix [ F(i,j) = incidence * w * prRecipMat(i,j), with w = beta_NH * beta_age * beta_care * beta_risk ]


prStageRecipMat:
  - Prob for EHI recipient to jump to next other CD4 stage
